---
title: "skhillon Models Creation"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)  # Misc tidy data wrapper package.
library(tools)  # Only used for toTitleCase()
library(class)  # KNN model  
library(ggplot2)  # Plotting
```

Load visas data
```{r}
load("visa_info.RData")
```

Convert all SOC, and Job Title fields to Title Case.
We leave Employers as all-caps because Company names vary greatly.
```{r}
visas$SOC_NAME <- toTitleCase(tolower(visas$SOC_NAME))
visas$JOB_TITLE <- toTitleCase(tolower(visas$JOB_TITLE))
```

## Classification - Case Status

KNN model to predict categorical Case Status given user input.
Since KNNs do not continuously "learn" and dimensionality is low, it is a useful choice.
Therefore, we also don't need to split data into training and testing because we will not be changing the number of clusters.
Source: https://dzone.com/refcardz/machine-learning-predictive?chapter=9
```{r}
num_cases <- unique(visas$CASE_STATUS)
set.seed(37)
visa_colnames <- colnames(visas)[3:11]
cluster_eval_cols <- setdiff(visa_colnames, "PREVAILING_WAGE")
cluster_eval_cols
str(visas[, cluster_eval_cols])
```

Not sure whether it wants numbers or factors, so here's both. (update: both fail)
```{r}
encode_numeric <- function(df) {
    df$emp_encoding <- as.numeric(as.factor(df$EMPLOYER_NAME))
    df$soc_encoding <- as.numeric(as.factor(df$SOC_NAME))
    df$job_title_encoding <- as.numeric(as.factor(df$JOB_TITLE))
    df$ft_encoding <- as.numeric(as.factor(df$FULL_TIME_POSITION))
    df$worksite_encoding <- as.numeric(as.factor(df$WORKSITE))
    df$year_encoding <- as.numeric(df$YEAR)
    
    return(df)
}

encode_factor <- function(df) {
    df$lon_encoding <- as.factor(df$lon)
    df$lat_encoding <- as.factor(df$lat)
    df$emp_encoding <- as.factor(df$EMPLOYER_NAME)
    df$soc_encoding <- as.factor(df$SOC_NAME)
    df$job_title_encoding <- as.factor(df$JOB_TITLE)
    df$ft_encoding <- as.factor(df$FULL_TIME_POSITION)
    df$worksite_encoding <- as.factor(df$WORKSITE)
    df$year_encoding <- as.factor(df$YEAR)
    
    return(df)
}

visas <- encode_numeric(visas)
head(visas)
```

Train-Test split, 70-30. Train = TRUE, Test = 2.
```{r}
v.indices <- sample(2, nrow(visas), replace = TRUE, prob = c(0.70, 0.30))
v.train <- visas[v.indices == 1, union(2, 10:17)]
v.test <- visas[v.indices == 2, union(2, 10:17)]

head(v.test)
```

Build classifier.
*Error: Error in sprintf(gettext(fmt, domain = domain), ...) : invalid format '%d'; use format %s for character objects*
```{r}
classifier <- knn(v.train, v.test, v.train$CASE_STATUS, k = num_cases)
classifier
```

```{r}
#predict_case <- function(user_df) {
#    user_numeric_df <- encode_numeric(user_df)
#    predict(kmeans_cluster, newdata = user_df[, cluster_eval_cols])
#}
```

## Regression - Yearly Wage Estimation (TODO)
```{r}
#wage_model <- lm()
```

## TODO: Output functions to RData so Shiny can load them.
