---
title: "skhillon Models Creation"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)  # Misc tidy data wrapper package.
library(tools)  # Only used for toTitleCase()
library(caret)  # KNN model  
library(ggplot2)  # Plotting
```

Load visas data
```{r}
load("visa_info.RData")
```

Convert all SOC, and Job Title fields to Title Case.
We leave Employers as all-caps because Company names vary greatly.
```{r}
visas$SOC_NAME <- toTitleCase(tolower(visas$SOC_NAME))
visas$JOB_TITLE <- toTitleCase(tolower(visas$JOB_TITLE))
```

## Classification - Case Status

KNN model to predict categorical Case Status given user input.
Since KNNs do not continuously "learn" and dimensionality is low, it is a useful choice.
Therefore, we also don't need to split data into training and testing because we will not be changing the number of clusters.
Source: https://dzone.com/refcardz/machine-learning-predictive?chapter=9
```{r}
num_cases <- length(unique(visas$CASE_STATUS))
set.seed(37)
visa_colnames <- colnames(visas)[3:11]
cluster_eval_cols <- setdiff(visa_colnames, "PREVAILING_WAGE")
cluster_eval_cols
str(visas[, cluster_eval_cols])
```

Not sure whether it wants numbers or factors, so here's both. (update: both fail)
```{r}
encode_numeric <- function(df) {
    df$emp_encoding <- as.numeric(as.factor(df$EMPLOYER_NAME))
    df$soc_encoding <- as.numeric(as.factor(df$SOC_NAME))
    df$job_title_encoding <- as.numeric(as.factor(df$JOB_TITLE))
    df$ft_encoding <- as.numeric(as.factor(df$FULL_TIME_POSITION))
    df$worksite_encoding <- as.numeric(as.factor(df$WORKSITE))
    df$year_encoding <- as.numeric(df$YEAR)
    
    return(df)
}

encode_factor <- function(df) {
    df$lon_encoding <- as.factor(df$lon)
    df$lat_encoding <- as.factor(df$lat)
    df$emp_encoding <- as.factor(df$EMPLOYER_NAME)
    df$soc_encoding <- as.factor(df$SOC_NAME)
    df$job_title_encoding <- as.factor(df$JOB_TITLE)
    df$ft_encoding <- as.factor(df$FULL_TIME_POSITION)
    df$worksite_encoding <- as.factor(df$WORKSITE)
    df$year_encoding <- as.factor(df$YEAR)
    
    return(df)
}

visas <- encode_numeric(visas)
head(visas)
```

Train-Test split, 70-30. Train = TRUE, Test = 2.
```{r}
in.train.rows <- createDataPartition(visas$CASE_STATUS, p = 0.7, list = FALSE)
in.train.cols <- union(2, 10:17)
v.train <- visas[in.train.rows, in.train.cols]
v.test <- visas[-in.train.rows, in.train.cols]

v.train$CASE_STATUS <- factor(v.train$CASE_STATUS)
head(v.train)
```

Build classifier.
*Error: Error in sprintf(gettext(fmt, domain = domain), ...) : invalid format '%d'; use format %s for character objects*
```{r}
train_control <- trainControl(method = "repeatedcv", number = 10, repeats = 3)
classifier <- train(CASE_STATUS ~., data = v.train, method = "knn",
                    trControl = train_control, 
                    preProcess = c("center", "scale"),
                    tuneLength = 10)
classifier
```

```{r}
predict_case_status <- function(user_df) {
    user_numeric_df <- encode_numeric(user_df)
    predict(classifier, newdata = user_numeric_df)
}
```

## Regression - Yearly Wage Estimation (TODO)
```{r}
wage_regressor <- lm(PREVAILING_WAGE ~ lon + lat + emp_encoding + soc_encoding 
                     + job_title_encoding + ft_encoding + worksite_encoding + year_encoding,
                 data = visas)
wage_regressor
```

```{r}
predict_wage <- function(user_df) {
    user_numeric_df <- encode_numeric(user_df)
    predict(wage_regressor, newdata = user_numeric_df)
}
```

```{r}
summary(wage_regressor)
```
