---
title: "Visa Exploration - Part 2"
output:
  html_document:
    df_print: paged
---

This document contains visualizations for part 2 of the STAT 331 final project.
```{r}
library(tidyverse)
library(readr)
library(stringr)
library(leaflet)
```

Because the data is so large, we will only import the first 10,000 observations:
```{r}
h1b.apps <- read.csv("~/Desktop/DataScienceCourses/stat331/ShinyVisaExploration/h1b_kaggle.csv", nrows = 10000)
str(h1b.apps)
```


## Geographic Exploration

The code below generates a leaflet plot with frequency of applications. In an interactive setting, this map could include filters to view by:

- Year
- Application status
- Employer name
- SOC name
- Job Title
- Full Time/part time
- Wage

Below is the data without any filters:
```{r}
leaflet(h1b.apps) %>% 
    addTiles() %>%
    addMarkers(~lon, ~lat, clusterOptions = markerClusterOptions())
```

Below is an example of the same map with the following filters:

- Application Certified and not withdrawn
- Full time jobs
- Year 2011, 2013, or 2016
- Is some sort of Executive or Physicist

```{r}
sample.map.filtered <- h1b.apps %>%
    filter(CASE_STATUS == "CERTIFIED" &
               FULL_TIME_POSITION == "Y" & 
               (YEAR == 2011 | YEAR == 2013 | YEAR == 2016) & 
               PREVAILING_WAGE > 100000 & 
               grepl("executive|physicists", SOC_NAME, ignore.case = TRUE))

leaflet(sample.map.filtered) %>% 
    addTiles() %>%
    addMarkers(~lon, ~lat, clusterOptions = markerClusterOptions())
```


